---
title: "Estimating transmissibility with population stratification"
author: Thibaut Jombart, Hugo Gruson
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document
link-citations: no
params:
  epicurve_unit: "week"
  incomplete_days: 7
  r_estim_window: 21
  si_mean: 4.2
  si_sd: 4.9
# bibliography: "../biblio/biblio.bib"
---


```{r settings, echo = FALSE}
knitr::opts_chunk$set(
  fig.width = 9,
  fig.height = 5,
  dpi = 90,
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  out.width = "100%",
  echo = FALSE
)
```

```{r}
library("tidyverse")
library("here")
library("rio")
library("linelist")
library("scales")
library("janitor")
library("rmdformats")
library("kableExtra")
library("incidence2")
library("i2extras")
library("distcrete")
library("epitrix")
```

```{r}
custom_grey <- "#505B5B"
green_grey <- "#5E7E80"
pale_green <- "#b2d1cc"
dark_green <- "#005C5D"
dark_pink <- "#B45D75"

theme_custom <- theme(
  plot.background = element_rect(fill = "white"),
  panel.grid.major = element_line(colour = "#96a3a3"),
  panel.grid.minor = element_line(colour = "#9DC1C2", linetype = "dotted"),
  panel.background = element_rect(colour = custom_grey, fill = "white"),
  strip.background = element_rect(fill = green_grey, color = "white", size = 1),
  strip.text = element_text(colour = "white"),
  text = element_text(size = 16, colour = custom_grey),
  axis.ticks = element_line(colour = custom_grey)
  #aspect.ratio = 9 / 16
)
theme_set(theme_custom)
```

```{r}
dat_raw <- "https://github.com/epiverse-trace/data_pipelines/raw/main/data/covid_admissions_uk_2020_10_24.xlsx" %>%
  rio::import() %>%
  tibble() %>%
  mutate(date = as.Date(date)) # date-time -> date
```

```{r}
date_var <- "date"
group_var <- "region"
count_var <- "n"

dat <- dat_raw %>%
  make_linelist(date_admission = date_var,
                location = group_var,
                counts = count_var,
                allow_extra = TRUE)
```

# Descriptive analyses

## Epidemic curves

This section creates epidemic curves ("_epicurves_"), with or without stratification.

```{r}
# convert daily incidence into weekly incidence using incidence2
dat_i <- dat_raw %>%
  incidence("date",
            interval = params$epicurve_unit,
            counts = {{count_var}},
            groups = {{group_var}})

# general variables for automatic customisation of plots
n_groups <- length(unique(unlist(dat_i[group_var])))
small_counts <- max(dat_i[count_var]) < 20

# plot with regions as colors if less than 9 categories
if (n_groups <= 9L) {
dat_i %>%
  plot(fill = group_var, col_pal = muted, color = "white",
       alpha = 1, angle = 45, legend = "bottom",
       show_cases = small_counts) +
  theme_custom +
  guides(fill = guide_legend(title=NULL)) +
  labs(title = "Incidence of cases over time")
} else {
dat_i %>%
  regroup() %>% 
  plot(angle = 45, color = "white",
       fill = green_grey,
       alpha = 1,
       show_cases = small_counts) +
  labs(title = "Incidence of cases over time") +
  theme_custom
}
```

```{r fig.height = 5 / 3 * n_groups}
dat_i %>%
  facet_plot(angle = 45,
             date_format = "%d %b %y",
             color = "white",
             fill = green_grey,
             alpha = 1,
             n_breaks = 6,
             nrow = n_groups,
             show_cases = small_counts) +
  theme_custom
```

## Numbers of cases

This graph shows the total number of cases per group:

```{r }
total_cases <- dat %>%
  select_tags(location, counts) %>%
  group_by(location) %>%
  summarise(cases = sum(counts)) %>%
  arrange(location)

ggplot(total_cases, aes(x = cases, y = location)) +
  geom_col(fill = green_grey) +
  theme_custom + 
  labs(x = "Total number of cases", y = NULL)

total_cases %>%
  mutate(percentage = cases / sum(cases),
         percentage = paste(round(percentage * 100, 2), "%")) %>% 
  adorn_totals() %>%
  mutate(cases = format(cases, scientific = FALSE, big.mark = " ")) %>% 
  set_names(toupper) %>%   
  kbl() %>%
  kable_paper("striped", font_size = 18, full_width = FALSE) %>%
  row_spec((1:nrow(total_cases))[c(FALSE, TRUE)],
           background = pale_green)
```

*************************
# Serial interval distribution

## Explanations

The _serial interval_ ($si$) is the delay between the date of symptom onsets of primary
case and the secondary cases they have infected. Because this delay varies from
one transmission pair to another, we will characterise this variation using a
probability distribution. This distribution is a key input to methods use for
estimating the reproduction number ($R$). 

Here, we assume that the mean and standard deviation of the $si$ is known, and
provided as an input by the user. We model the $si$ distribution as a
discretized Gamma. 

## Results

```{r}
si_mean <- params$si_mean
si_sd <- params$si_sd
si_cv <- si_sd / si_mean
si_params <- epitrix::gamma_mucv2shapescale(mu = si_mean, cv = si_cv)
si <- distcrete("gamma",
                interval = 1,
                w = 0,
                shape = si_params$shape,
                scale = si_params$scale)
si_x <- seq(1L, to = si$q(.999), by = 1L)

ggplot(data.frame(delay = si_x, prob = si$d(si_x)),
       aes(x = delay, y = prob)) +
  geom_col(fill = green_grey) +
  labs(title = "Serial interval distribution",
       x = "Days from primary to secondary onset",
       y = "Probability",
       subtitle = sprintf(
         "Gamma distribution | mean: %.1f days ; sd: %.1f days",
         si_mean, si_sd))
```

# Growth rates and doubling/halving times

## Explanations

The package *i2extras* is used to estimate daily growth rates '_r_', using Poisson
GLMs fitted to epidemic curves. These estimates can be biased by reporting
delays. Here, we exclude the last `r params$incomplete_days` days as incomplete,
and retain a total of `r params$r_estim_window` days to estimate the growth
rate.

Growth rates can be interpreted as the percent change in daily
incidence. Positive numbers indicate growth, while negative numbers indicate
decline, of the epidemic. For instance:

* $r = 0.02$ means that the number of new cases increases on average by 2% each day
* $r = -0.13$ means that the number of new cases decreases on average by 13% each day

Growth rates can also be converted to _doubling times_ (when $r > 0$) or
_halving times_ (when $r < 0$), defined as the time it takes for the daily
incidence to double (respectively, halve).

For more information on growth rates, see this [blog post by Pr Julia Gog](https://plus.maths.org/content/epidemic-growth-rate).

## Results

This section contains:

* graphs showing the models fitted to epidemic curves, with their associated 95% confidence
intervals (CI)

* graphs showing the estimates of the growth rates ($r$) with their 95% CI

* a table summarizing estimtes of $r$ and the associated period (doubling or
  halving time); all results show point estimates and their 95% CI

```{r fig.height = 5 / 3 * n_groups}
last_date <- dat %>%
  pull(date) %>%
  max()

# version using keep_first and keep_last from i2extras
days_to_keep <- params$incomplete_days + params$r_estim_window
i_recent <- dat_raw %>%
  incidence("date",
            counts = {{count_var}},
            groups = {{group_var}}) %>% 
  keep_last(days_to_keep) %>%  # keep data for fitting 
  keep_first(params$r_estim_window) # remove incomplete data

last_trends <- i_recent %>%
  fit_curve(model = "poisson")
# version with negbin model needs more iterations to converge
# fit_curve(model = "negbin", control = glm.control(maxit = 1e3))
plot(last_trends,
     angle = 45,
     date_format = "%d %b %y",
     color = "white",
     fill = green_grey,
     alpha = .8,
     n_breaks = 12,
     nrow = n_groups,
     show_cases = small_counts)
```

```{r}
last_trends %>%
  growth_rate %>%
  ggplot(aes(y = .data[[group_var]]), fill = custom_grey) +
  geom_point(aes(x = r), color = dark_green) +
  geom_errorbar(aes(xmin = r_lower, xmax = r_upper), color = dark_green) +
  theme_custom +
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = 0, color = dark_pink) +
  labs(title = "Estimates of daily growth rates",
       subtitle = sprintf(
         "based on data from %s - %s",
         format(min(get_dates(i_recent)), "%d %B %Y"),
         format(max(get_dates(i_recent)), "%d %B %Y")),
       y = "",
       x = "Daily rate of change")

last_trends %>%
  growth_rate() %>%
  dplyr::select(-c(1,3)) %>%
  rename(period = growth_or_decay) %>% 
  mutate("period type" = paste(period, "time"),
         r = paste(round(r * 100, 1), "%"),
         r_ci = sprintf(
           "[%1.1f %% ; %1.1f %% ]",
           r_lower * 100,
           r_upper * 100),
         time = round(time, 1),
         time_ci = sprintf(
           "[%.1f ; %.1f]",
           time_lower,
           time_upper)
         ) %>%
  dplyr::select("Daily growth rate" = r,
         "Growth rate (95% CI)" = r_ci,
         "period type",
         "Duration (days)" = time,
         "Duration (95% CI)" = time_ci) %>%
  set_names(toupper) %>%   
  kbl() %>%
  kable_paper("striped", font_size = 18, full_width = FALSE) %>%
  row_spec((1:nrow(total_cases))[c(FALSE, TRUE)],
           background = pale_green)
```

# Estimating $R$ from the growth rates

## Explanations

Wallinga and Lipsitch have introduced methods for converting daily growth rates
($r$) into reproduction numbers ($R$) [@Wallinga2007-hw], when information on
the serial interval distribution is available. $R$ represents the average number
of secondary cases generated by an infected patient. Like the daily growth rate,
$R$ can be used to gauge the dynamics of an epidemic:

* $R > 1$: the epidemic will grow exponentially
* $R < 1$: the epidemic will decline exponentially

Note that $R$ only speaks to the number of new cases created, not how fast these
are generated, so on its own it is not sufficient to indicate how fast the
epidemic is growing. For a comparison of $r$ and $R$, we recommend reading this
[blog post by Pr Julia Gog](https://plus.maths.org/content/epidemic-growth-rate).

Here, we use the procedure implemented in `epitrix` to derive samples of $R$
values compatible with the values of $r$ estimated in the previous section, and
with the provided serial interval.

## Results

This section contains:

* a graph showing the distribution of $R$ values estimated from the daily growth
  rates; each violin shows the density of values based on a sample of 500 from
  the Student distribution associated with the estimate of $r$; the line range
  represents the 95% confidence interval (horizontal segment) and the median
  (dot)
  
* a table summarizing the estimated distributions of $R$ values 

```{r }
res_R_wl <- last_trends %>%
  mutate(R = map(model, epitrix::lm2R0_sample, w = si, 500)) %>%
  dplyr::select({{group_var}}, R) %>%
  unnest(R)

res_R_wl_smry <- res_R_wl %>%
  group_by(.data[[group_var]]) %>%
  summarise(mean = mean(R),
            median = median(R),
            q_0.025 = quantile(R, 0.025),
            q_0.975 = quantile(R, 0.975))

res_R_wl %>%
  ggplot(aes(y = .data[[group_var]]), fill = custom_grey) +
  geom_violin(aes(x = R), fill = pale_green, color = green_grey) +
  geom_pointrange(data = res_R_wl_smry,
                  aes(x = median, xmin = q_0.025, xmax = q_0.975),
                  color = dark_green) +
  geom_vline(xintercept = 1, color = dark_pink) +
  theme_custom +
  labs(title = "Reproduction numbers estimated from growth rates",
       subtitle = sprintf(
         "based on data from %s - %s",
         format(min(get_dates(i_recent)), "%d %B %Y"),
         format(max(get_dates(i_recent)), "%d %B %Y")),
       y = "",
       x = "Reproduction number")

res_R_wl_smry %>%
  mutate(
    mean = round(mean, 2),
    median = round(median, 2),
    `95% ci` = sprintf(
           "[%1.2f ; %1.2f]",
           q_0.025,
           q_0.975)) %>%
  dplyr::select(-c(q_0.025, q_0.975)) %>%
  rename(
    "mean $R$" = mean,
    "median $R$" = median) %>% 
  set_names(toupper) %>%   
  kbl() %>%
  kable_paper("striped", font_size = 18, full_width = FALSE) %>%
  row_spec((1:nrow(res_R_wl_smry))[c(FALSE, TRUE)],
           background = pale_green)
```

# Estimating $R_t$ using _EpiEstim_ 

## Explanations


## Results

### Global transmissibility

These analyses present results for the global incidence, i.e. without
stratification. Results include:
 
* a graph of the estimates of $R_t$ over time, with associated 95% credibility
  intervals
  
* a table summarising these results for the last two weeks


```{r }
library(EpiEstim)

# Function to process a distrcrete object and output `si_discr` argument for
# EpiEstim::make_config

wrap_si <- function(x) {
  stopifnot(inherits(x, "distcrete"))
  max_x <- x$q(.999)
  x <- si$d(seq_len(max_x))
  x[1] <- 0
  x <- x / sum(x)
  x  
}


# Function to process output from `EpiEstim::estimate_R` and output a tibble
# with dates, mean R, and associated confidence intervals
#'
#' @param x An `estimate_R` object
#' @param incid (Optional) The incidence2 object on which R values have been
#'   estimated.
wrap_res <- function(x, incid = NULL) {
  stopifnot(inherits(x, "estimate_R"))
  out <- tibble::tibble(x$R)
  out <- dplyr::select(out,
                       start = t_start,
                       end = t_end,
                       mean = `Mean(R)`,
                       sd = `Std(R)`,
                       median = `Median(R)`,
                       lower = `Quantile.0.025(R)`,
                       upper = `Quantile.0.975(R)`)
  if (!is.null(incid)) {
    stopifnot(inherits(incid, "incidence2"))
    dates <- unique(incidence2::get_dates(incid))
    out$start <- dates[out$start]
    out$end <- dates[out$end]
  }
  
  class(out) <- c("R_estimate", class(out))
  out
}

# Daily incidence
dat_i_day <- dat_raw %>%
  incidence("date",
            interval = 1L,
            counts = {{count_var}},
            groups = {{group_var}})

# Run EpiEstim
ee_config <- make_config(si_distr = wrap_si(si))

res_epiestim_global <- dat_i_day %>% 
  regroup() %>%
  pull(n) %>%
  estimate_R(config = ee_config) %>%
  wrap_res(dat_i_day)

# Graph of all values over time
ggplot(res_epiestim_global, aes(x = end)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = pale_green) +
  geom_line(aes(y = median), color = dark_green) +
  geom_hline(yintercept = 1, color = dark_pink) +
  scale_x_date(breaks = scales::pretty_breaks(n = 8), date_label = "%d %b %Y") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_custom +
  labs(x = "",
       y = "Instantaneous reproduction number\\(Rt)",
       title = "Estimates of Rt (EpiEstim)")

# Table
res_epiestim_global %>%
  tail(14) %>% 
  mutate(
    date = end,
    mean = round(mean, 2),
    median = round(median, 2),
    `95% ci` = sprintf(
           "[%1.2f ; %1.2f]",
           lower,
           upper)) %>%
  dplyr::select(-c(start, end, sd, lower, upper)) %>%
  dplyr::select(date, everything()) %>%
  rename(
    "mean $R$" = mean,
    "median $R$" = median) %>% 
  set_names(toupper) %>%   
  kbl() %>%
  kable_paper("striped", font_size = 18, full_width = FALSE) %>%
  row_spec((1:14)[c(FALSE, TRUE)],
           background = pale_green)
```



### Transmissibility by group

```{r fig.height = 5 / 3 * n_groups}
# Get results by group, keeping only the last 14 days of data
res_epiestim_group <- dat_i_day %>% 
  nest(data = c(date_index, n)) %>%
  mutate(res_epiestim = map(data, ~ wrap_res(estimate_R(.$n, config = ee_config), dat_i_day))) %>%
  unnest(res_epiestim) %>%
  dplyr::select(-data)

# Graph of all values over time
ggplot(res_epiestim_group, aes(x = end)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = pale_green) +
  geom_line(aes(y = median), color = dark_green) +
  geom_hline(yintercept = 1, color = dark_pink) +
  scale_x_date(breaks = scales::pretty_breaks(n = 8), date_label = "%d %b %Y") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_custom +
  facet_wrap( ~ .data[[group_var]], ncol = 1) +
  labs(x = "",
       y = "Instantaneous reproduction number\\(Rt)",
       title = "Estimates of Rt (EpiEstim)")
```

```{r }
# Plot of latest estimates
res_epiestim_group %>%
  filter(end == max(end)) %>%
  ggplot(aes(y = .data[[group_var]]), fill = custom_grey) +
  geom_point(aes(x = median), color = dark_green) +
  geom_errorbar(aes(xmin = lower, xmax = upper), color = dark_green) +
  geom_vline(xintercept = 1, color = dark_pink) +
  theme_custom +
  labs(title = "Estimates of Rt (EpiEstim)",
       subtitle = sprintf(
         "based on data from %s - %s",
         format(max(get_dates(dat_i_day) - 7), "%d %B %Y"),
         format(max(get_dates(dat_i_day)), "%d %B %Y")),
       y = "",
       x = "Instantaneous Reproduction Number (Rt)")

res_epiestim_group %>%
  filter(end == max(end)) %>% 
  mutate(
    mean = round(mean, 2),
    median = round(median, 2),
    `95% ci` = sprintf(
           "[%1.2f ; %1.2f]",
           lower,
           upper)) %>%
  dplyr::select(-c(sd, lower, upper)) %>%
  rename(
    "mean $R$" = mean,
    "median $R$" = median) %>% 
  set_names(toupper) %>%   
  kbl() %>%
  kable_paper("striped", font_size = 18, full_width = FALSE) %>%
  row_spec((1:n_groups)[c(FALSE, TRUE)],
           background = pale_green)
```



*************************
# Estimating $R_t$ using _R0_

## Explanations

## Results

```{r}
library(R0)

mGT <- generation.time("gamma", c(3, 1.5))

# Helper function to get R0 outputs in a tidy format, compatible with pipelines
r0_quantiles <- function(Rt) {
  
  data.frame(
    q50   = Rt$R,
    q2.5  = Rt$conf.int$lower,
    q97.5 = Rt$conf.int$upper
  )
  
}

last_trends_r0 <- dat %>% 
  filter(between(.data[[date_var]], last_date - (params$incomplete_days + params$r_estim_window), last_date - params$incomplete_days)) %>% 
  arrange(.data[[date_var]]) %>% 
  group_by(.data[[group_var]], .data[[date_var]]) %>% 
  summarise({{count_var}} := sum(.data[[count_var]])) %>% 
  group_by(.data[[group_var]]) %>% 
  summarise(
    r0_quantiles(estimate.R(.data[[count_var]], mGT, .data[[date_var]], begin = 1L, end = params$r_estim_window, nsim = 1e4, method = "TD")$estimates$TD),
    {{date_var}} := head(.data[[date_var]], params$r_estim_window)
  )

last_trends_r0 %>% 
  ggplot(aes(x = date, y = q50, ymin = `q2.5`, ymax = `q97.5`)) +
    geom_ribbon(fill = "grey70") +
    geom_line() +
    facet_wrap(vars(region)) +
    geom_hline(yintercept = 1, color = "red", linetype = "dotted") +
    labs(title = "Estimates of effective reproduction numbers (Reff)",
       subtitle = sprintf(
         "based on data from %s - %s",
         format(last_date - (params$incomplete_days + params$r_estim_window), "%d %B %Y"),
         format(last_date - params$incomplete_days, "%d %B %Y")),
       y = "Reff") +
    theme_custom
```



*************************
# Estimating $R_t$ using _EpiNow2_ 

## Explanations

## Results




